---
title: "Nordic Ski Analysis"
author: "Daniel Muzyka"
date: "2/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries

```{r, libraries, message=FALSE}
library(tidyverse)
library(hexbin)
```

## Reading in Data

Pulling my data straight from Github and filtering for Nordic Skiing data

```{r, read in data}

strava <- readRDS(gzcon(url("https://github.com/muzykad397/strava/blob/main/data/strava.RDS?raw=true")))

#this is an analysis for nordic skiing only so I will filter the data now
nordic <- strava %>% filter(type=="NordicSki") 
```

# Basic Plots

### Histograms

##### Looking at distributions
We can see a a curve starting to take shape. As of writing I only have 36 observations. Perhaps I will have a better curve by the end of the season. I see a pesky grouping of short trips on the left side. These are certainly quick jaunts on the trail by my house when we received fresh snow.

```{r basic plot, echo=FALSE}
#Histogram of distance
hist(nordic$distance, breaks = 21, main = "Histogram of Nordic Ski Distances", xlab = 'Distance (Miles)')

#Histogram of elevation gain
hist(nordic$total_elevation_gain, breaks = 21, main = "Histogram of Total Elevation Gains", xlab = 'Elevation Gains (Feet)')

#Histogram of distance by season
nordic %>% ggplot() + geom_histogram( aes(x=distance, fill=season),binwidth = .5)

#Histogram of elevation gain
nordic %>% ggplot() + geom_histogram( aes(x=total_elevation_gain, fill=season),binwidth = 50)
```

## Looking at relationships

### Scatter Plots

We can see in general there is a relationship between distance and total elevation gain.
This overall is not surprising because my main venue of Mt. Spokane features rolling hills. 

```{r, scatters}
#scatter plot of elevation gain * distance with a linear model fit at .05 alpha 
nordic %>% ggplot() + geom_point(aes(x=total_elevation_gain, y=distance, color=season)) + 
  geom_smooth(aes(x=total_elevation_gain, y=distance),method=lm, formula=y~x, level=0.95)

#scatter plot of elevation gain * distance with a loess fit at .05 alpha
nordic %>% ggplot() + geom_point(aes(x=total_elevation_gain, y=distance, color=season)) + 
  geom_smooth(aes(x=total_elevation_gain, y=distance),method=loess,formula=y~x, level=0.95)
```

### Hexplot
In the hexplot I am starting to see a frequency in the center showing my typical ski is around 6 miles and 750 feet of elevation gain. 
```{r, hexplot}
#Hexbin to show not only relationshp, but frequency of the points
nordic %>% ggplot(aes(x = total_elevation_gain, y = distance)) + 
  geom_hex() +
  scale_fill_viridis_c()

```


### Barchart

Here we can see nice distributions for my distance traveled and elevation climbed per month.

I wonder if there is a third variable involved...***
```{r, barchart by month}
#reorder month for a nice curve through the winter months
nordic$month = with(nordic, reorder(month, month_order))

nordic  %>%  ggplot() + geom_col( aes(x=month,y=distance,fill=season))

nordic  %>%  ggplot() + geom_col( aes(x=month,y=total_elevation_gain,fill=season))
```

## Creating seasonal summary data

I would like to plot my season totals in a cumulative fashion. This will help me see if I am getting more skiing done compared to last year around this time.

This step is creating the aggregated data

```{r data prep for mileage season totals}
#I still need to add cumulative totals. 
#Before I do this I want to fill any months without data with a 0 mileage for illustration purposes
#for example, one season starts earlier, or in one season I happen to not ski an entire month 
min<-nordic %>%  summarise(.,min(month_order))
max<-nordic %>%  summarise(.,max(month_order))
seasons <- nordic %>% distinct(.,season)
month_translate <- nordic %>% group_by(month,month_order) %>% distinct(month,month_order) %>% ungroup()

min_to_max <- data.frame(month_order=min$'min(month_order)':max$'max(month_order)',
                        distance=0,
                        total_elevation_gain=0)  %>% 
  full_join(.,month_translate, by="month_order") %>% full_join(.,seasons,by=character())



summaries <- bind_rows(nordic,min_to_max) %>% group_by(season,month,month_order) %>% 
  summarise(mileage = sum(distance),
            elevation_gain = sum(total_elevation_gain),
                              .groups = 'drop') %>% 
  arrange(season,month_order) %>% group_by(season) %>% 
  mutate(.,cum_mileage=cumsum(mileage)) %>% 
  mutate(.,cum_elevation=cumsum(elevation_gain))


summaries$month = with(summaries, reorder(month, month_order))
```

## Plotting Season to Date totals

Now that I have summarized data, I would like to plot it.

We can see that as of today I am neck-in-neck with last season's totals. The good new is this is summarized by month and today is only the 10th of February!

```{r plooting season totals}

summaries %>% 
ggplot(.,aes(x=month)) + ggtitle("Cumulative X-Country Skiing Mileage") +
   theme(plot.title = element_text(hjust = 0.5)) +
  geom_bar( aes(y=mileage,group=season,fill=season), stat="identity" , position="dodge") +
 geom_line( aes(y=cum_mileage, group=season, color=season),stat="identity", size=1) +
  
  scale_y_continuous(
    name="Miles per Month (Bars)",
    sec.axis = sec_axis(~.,name="Cumulative Miles per Season (Lines)")
    
  ) 

summaries %>% 
ggplot(.,aes(x=month)) + ggtitle("Cumulative X-Country Elevation Gain") +
   theme(plot.title = element_text(hjust = 0.5)) +
  geom_bar( aes(y=elevation_gain,group=season,fill=season), stat="identity" , position="dodge") +
 geom_line( aes(y=cum_elevation, group=season, color=season),stat="identity", size=1) +
  
  scale_y_continuous(
    name="Total Elevation Gain per Month (Bars)",
    sec.axis = sec_axis(~.,name="Cumulative Elevation Gain per Season (Lines)")
    
  ) 

```