---
title: "Ski_Versus_Board"
author: "Daniel Muzyka"
date: "4/14/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Skiing Versus Snowboarding

After a season of doing the leg work to collect data, I want to answer the question, "Am I a skier or a snowboarder"?

## Loading Libraries

```{r, libraries, message=FALSE}
library(tidyverse)
library(lubridate)
```

## Import

```{r import}
strava <- readRDS(gzcon(url("https://github.com/muzykad397/strava/blob/main/data/strava.RDS?raw=true")))

#set the list of downhill wintersports activities I want to compare.

dh_activities <- c("AlpineSki","Snowboard")

dh_sports <- strava %>% filter(., type %in% dh_activities) %>% mutate(., date=date(start_date))
```

## Limiting to 21/22 season

There was one day where strava did not sync my data so I did a manual entry. I will remove it here.

```{r}
dh_sports <- dh_sports %>% filter(., season == "2021/2022 season") %>% filter (., max_speed != 0)
```
## Basic summaries

```{r summaries}
#days spent on each sport
  dh_sports %>% 
      group_by(season,type) %>% summarize(number_of_days_by_activity = n_distinct(date),.groups = 'drop')

#distance traveled for each sport
  dh_sports %>% 
      group_by(season,type) %>% summarize(total_season_distance_miles = sum(distance),.groups = 'drop')
  
#Total moving time spent on each sport
  dh_sports %>% 
      group_by(season,type) %>% summarize(time_spent_moving_hours = (sum(moving_time)/60**2),.groups = 'drop')
  
#Average distance traveled per day of activity
  dh_sports %>% 
      group_by(season,type) %>% summarize(avg_daily_distance_miles = mean(distance),.groups = 'drop')

#Average time spent per day
  dh_sports %>% 
      group_by(season,type) %>% summarize(avg_daily_time_spent_moving_hours = (mean(moving_time)/60**2),.groups = 'drop')
  
#Average of the average daily speed
  dh_sports %>% 
      group_by(season,type) %>% summarize(avg_avg_speed = mean(average_speed),.groups = 'drop')
  
#Average top speed of each day
  dh_sports %>% 
      group_by(season,type) %>% summarize(avg_max_speed = mean(max_speed),.groups = 'drop')
  
#Max top speed of the season
  dh_sports %>% 
      group_by(season,type) %>% summarize(max_max_speed = max(max_speed),.groups = 'drop')
  
```
## Max Speed Analysis

There are issues comparing a lot of these metrics. Some days are longer than others. These days took place on various mountains with different terrains and sizes. Some days had better conditions. Some days I rode with different people. Some days involved alcohol.

The only thing I can be sure of, is that everyday I had a need for speed. Regardless of what the overall day was like, there was at least one point, on one run, where I pointed my tip(s) downhill and let gravity do the work. So the value I want to perform statistical tests on is my daily max speed. 

```{r}
#simple box plot. One outlier snowboarding. I remember that exact run. It was fast.
dh_sports %>% ggplot(., aes(type,max_speed)) +
  geom_boxplot(aes(fill=type))

#labeling the observations with a rank variable so I can plot in different orders
obs= count(dh_sports)
dh_sports <- dh_sports %>% arrange(., max_speed) %>% mutate(., rank=1:obs$n)
dh_sports <- dh_sports  %>% arrange(., date) %>% mutate(., date_rank=1:obs$n)

#Bring out the means so I can plot as reference lines on the graphs
means <-   dh_sports %>% 
      group_by(season,type) %>% summarize(avg_max_speed = mean(max_speed),.groups = 'drop')

#plotted point in order of their ranking. 
#This might be fun to illustrate rank if I do a test like a Wilcoxon ranked sums
dh_sports %>% ggplot(., aes(rank,max_speed,type)) +
  geom_point(aes(color=type)) +
  geom_hline(data = means, aes(yintercept=avg_max_speed, color=type),
             group="type", linetype='dashed')

#This chart is looking at my speeds over time
#The idea behind this one is that I just started skiing this year. Maybe I will see improvement
#need to break into facets and change the reference line to show the slope of points
dh_sports %>% ggplot(., aes(date_rank,max_speed,type)) +
  geom_point(aes(color=type)) +
  geom_hline(data = means, aes(yintercept=avg_max_speed, color=type),
             group="type", linetype='dashed')

```

```{r, scatters}
#scatter plot of date_rank and max_speed with a linear fit
dh_sports %>% ggplot() + geom_point(aes(x=date_rank, y=max_speed, color=type, group=type)) + 
  geom_smooth(aes(x=date_rank, y=max_speed, group=type, color=type),method=lm, formula=y~x, level=0.3)

#scatter plot of date_rank and max_speed with a loess fit
dh_sports %>% ggplot() + geom_point(aes(x=date_rank, y=max_speed, color=type, group=type)) + 
  geom_smooth(aes(x=date_rank, y=max_speed, group=type, color=type),method=loess, formula=y~x, level=0.3)
```
