---
title: "Strava Data Import"
author: "Daniel Muzyka"
date: "2/10/2022"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Importing data from Strava into R with basic cleaning
#### Note, you must submit the code chunks through the import step before knitting. 
#### You must manually authorize API access via the terminal and a prompted webpage

Hello, this program is used to import data from the Strava API into R. If you do not already have a Strava API configured do a quick search on how to accomplish this. 

This program contains the code to access the Strava API, but I also included some of my cleaned personal Strava data as an output if you would like to play around first.

I have saved the outputs of this program to my repository:
  "muzykad397/strava/data"

## Libraries
These are the libraries required in the data import and cleaning steps.

```{r libraries, message=FALSE}
library(tidyverse)
library(httr)
library(yaml)
library(jsonlite)
library(janitor)
library(lubridate)
```

## The Import

I cannot claim the API import code as my own. This was adapted from various online sources.

**You must run up to this code chunk before knitting.** You will be required to authorize API access in the terminal.

**You will have to provide you own client_id and secret code.** These can be found by logging into Strava on the API management page.

If you do not have a Strava API, I have saved the output of this program to my repository:
  "muzykad397/strava/data"

```{r import}

client_id <- client_id
secret <- secret

app <- oauth_app("strava", client_id, secret)
endpoint <- oauth_endpoint(
  request = NULL,
  authorize = "https://www.strava.com/oauth/authorize",
  access = "https://www.strava.com/oauth/token"
)

token <- oauth2.0_token(endpoint, app, as_header = FALSE,
                        scope = "activity:read_all,activity:read,profile:read_all")

df_list <- list()
i <- 1
done <- FALSE
while (!done) {
  req <- GET(
    url = "https://www.strava.com/api/v3/athlete/activities",
    config = token,
    query = list(per_page = 200, page = i)
  )
  df_list[[i]] <- fromJSON(content(req, as = "text"), flatten = TRUE)
  if (length(content(req)) < 200) {
    done <- TRUE
  } else {
    i <- i + 1
  }
}

strava_raw <- rbind_pages(df_list)
```

## Basic cleaning of the data

```{r clean imported data}
strava <- strava_raw %>% clean_names() %>% 
  mutate(start_date = ymd_hms(start_date)) %>% 
  #I would like to convert all measurements from metric to imperial
    #distance in miles (meters ->miles)
      mutate(.,distance=distance*0.000621371) %>% 
    #elevation gain in feet (meters -> feet)
      mutate(.,total_elevation_gain=total_elevation_gain*3.28084) %>% 
    #elevation high in feet (meters -> feet)
      mutate(.,elev_high=elev_high*3.28084) %>% 
    #elevation low in feet (meters -> feet)
      mutate(.,elev_low=elev_low*3.28084) %>% 
    #average_speed in mph (m/s -> mph)
      mutate(.,average_speed=average_speed*2.23694) %>% 
    #max_speed in mph (m/s -> mph)
      mutate(.,max_speed=max_speed*2.23694) %>% 
  #Pull out month from start date
  mutate(.,month=format(.$start_date,"%m")) %>% 
  #the next bit of logic is to set a 'season' variable for my winter sports
    #this will be used for 'Season to date' calculations. 
    #Think 'year to date', but the winter season spans over the new year
    #I picked July as the cut off. Seems reasonable to me unless I find the funds to chase winter around the globe
  mutate(.,season = ifelse(
                            as.numeric(format(.$start_date,"%m")) <=7, 
        #if before July then current year is 2nd half of season
                                paste0(
                                  (as.numeric(format(.$start_date,"%Y"))-1),
                                   "/",
                                   format(.$start_date,"%Y"),
                                   " season"),
        #if after July then current year is 1st half of season
                                paste0(
                                  format(.$start_date,"%Y"),
                                   "/",
                                   (as.numeric(format(.$start_date,"%Y"))+1),
                                   " season")
                            )
            )  %>% 
  #month_order will be used later in graphing for the winter sports season
    #If I am doing season to date graphing I wont want order to be 1:12, I want 8:12,1:7
 mutate(.,month_order = ifelse(
                            as.numeric(month) <=7, 
        #if before July then current year is 2nd half of season
                                as.numeric(month)+5,
        #if after July then current year is 1st half of season
                               as.numeric(month)-7
                            )
            ) %>% 
  #To simplify reviewing the data I only kept variables of interest
    #For the output I save for public use I will drop the latitude and longitude
  select(.,distance:type, start_date,month, month_order, achievement_count,start_latlng:max_speed,elev_high, elev_low, pr_count,season)
```

## Saving data for later use (to be accessed if skipping API)

I have saved the outputs of this program to my repository:
  "muzykad397/strava/data"

```{r, save data}
#I will be dropping Latitude and Longitude data before posting publicly

#I will save as an R data set here
strava %>% select(., -start_latlng, -end_latlng
                  #, -start_latitude, -start_longitude
                  ) %>% 
    saveRDS(., file = "D:/Users/Danny/Documents/GitHub/strava/data/strava.RDS")

#I will save as csv for importing to other languages (like sas)
strava %>% select(.,-start_latlng, -end_latlng
                  #, -start_latitude,-start_longitude
                  ) %>% 
    write.csv(., file = "D:/Users/Danny/Documents/GitHub/strava/data/strava.csv")
```

