---
title: "Winter Sports Report"
author: "Daniel Muzyka"
date: "2/17/2022"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries

```{r, libraries, message=FALSE}
library(tidyverse)
library(lubridate)
```

## Reading in Data

Pulling my data straight from Github and filtering to specified winter sports activities.


```{r}
strava <- readRDS(gzcon(url("https://github.com/muzykad397/strava/blob/main/data/strava.RDS?raw=true")))

#set a list of wintersports activities I commonly use. Update to your needs.

winter_activities <- c("NordicSki","AlpineSki","Snowboard","Snowshoe")

winter_sports <- strava %>% filter(., type %in% winter_activities) %>% mutate(., date=date(start_date))

```

## High level Summary

How many days did I spend doing winter sports each season?

* disclaimer - I did not record all snowboarding trips in then 20/21 season)
  
#### Overall Winter Sports Days
```{r}
#overall number of days spent participating in winter sports by season
  winter_sports  %>%
      group_by(season) %>% summarize(winter_sport_days = n_distinct(date))
```

#### Days by Individual Sport

```{r}
#days spent on each type of winter sport
  winter_sports %>% 
      group_by(season,type) %>% summarize(number_of_day_by_activity = n_distinct(date),.groups = 'drop')
```

#### Count of Full Day Activity

Some days I participated in more than one sport. This summary reflects that.
```{r}
#some logic to roll up activities in a single day.
      #Example 1 - I Nordic ski and snowboard in the same day
      #Example 2 - I switch between Alpine Ski and Snowboarding in the same day
      #Example 3 - I recorded two Nordic Skis in one day (this is already accounted for in the summaries above)
      #Example 4 - I Snowshoe, Nordic Ski, and Snowboard all in one day! (will work but the category will have funky                     grammar)
combos <- winter_sports  %>%
  #sort order is imperative for lead+lag and also for binding_cols in a future step
  arrange(.,date) %>% 
  mutate(.,combo=case_when(
    #logic to create the combinations
    lead(date) == date & lead(type) != type ~ paste0(lead(type)," and ",type),
    lag(date) == date & lag(type) != type ~ paste0(lag(type)," and ",type),
    #some safety logic to account for Example 3 above
    lead(date) == date & lead(type) == type ~ type,
    lag(date) == date & lag(type) == type ~ type,
    TRUE ~ type
    ))

#Here I am making sure that my Combo days all follow the same format
  #If I ski then snowboard one day, and snowboard then ski on another, the lead+lag logic will create different          categories
  #To make sure things are conistent I will parse the combo field, sort the strings, then concatenate back together
 combos2 <-   str_split(combos$combo," and ") %>% lapply(.,str_sort) %>% 
      sapply(., paste, collapse = " and ")  %>% as.data.frame() %>% 
   transmute(.,full_day_activity_type= case_when(
    grepl("and",.) ~ paste0("Combo ",.),
    TRUE ~ .
   ))

 #bind the new column back on and roll up to full day summaries
  #make sure to preserve sort order from above because I am binding rows, not joining
 bind_cols(combos, combos2) %>% group_by(season,full_day_activity_type) %>% summarize(number_of_days = n_distinct(date),.groups = 'drop')


```